\documentclass{article}
\usepackage{dtrt}
\usepackage{graphicx} % Required for inserting images
\usepackage{fontspec}
\usepackage{libertine}
\usepackage{enumitem}
\usepackage{fullpage}
\usepackage{amsmath}
\usepackage{tikz}
\usepackage{hyperref}
\usetikzlibrary{arrows.meta, shapes.geometric, positioning, fit, backgrounds, calc}

% BibTeX configuration
% Format: BMS24 style (author initials + two-digit year, e.g., [BMS24])
\bibliographystyle{alpha}
\setlength{\parindent}{0pt} % No indentation for paragraphs
\setlength{\parskip}{0.5em} % Add gap between paragraphs

\title{Private Information Retrieval for Bitcoin Light Clients}
\author{Weikeng Chen}
\date{March 2026}

\begin{document}

\maketitle

\section{Introduction}

We are interested in using Private Information Retrieval (PIR) to help Bitcoin light clients synchronize with the Bitcoin network privately. A light client that knows its own keys needs to discover which unspent transaction outputs (UTXOs) it can spend, without revealing its addresses to the servers it queries. PIR enables this by allowing the client to retrieve specific database entries without the server learning which entries were accessed.

This paper presents a complete system for private UTXO lookups based on 2-server DPF-PIR with Probabilistic Batch Codes. Section~\ref{sec:background} provides the necessary background on PIR, distributed point functions, cuckoo hashing, batch codes, and the Bitcoin UTXO model. Section~\ref{sec:landscape} surveys the PIR protocol landscape across single-server and two-server settings. Section~\ref{sec:utxo-pir} describes our two-level architecture for private UTXO lookups. Section~\ref{sec:implementation} details the implementation, and Section~\ref{sec:extension} discusses UTXO set reduction techniques.

\clearpage

\section{Background}
\label{sec:background}

\subsection{Private Information Retrieval}

Private Information Retrieval (PIR) allows a client to retrieve an entry from a database held by one or more servers without revealing which entry is being accessed. Formally, consider a database $D$ of $N$ entries, each of $B$ bits. The client holds an index $i \in [N]$ and wishes to retrieve $D[i]$.

The trivial solution is for the client to download the entire database, achieving perfect privacy at $O(NB)$ communication cost. The goal of PIR is to achieve sublinear communication while preserving privacy.

\textbf{Single-server (computational) PIR.} A single server holds $D$. The client's query must be computationally indistinguishable from a query for any other index under some hardness assumption. The server performs computation over the query and the database to produce a response, from which the client can extract $D[i]$. The server's computation is at least $\Omega(NB)$ in the worst case (otherwise the server trivially learns that certain entries were not queried), though preprocessing and client-side hints can reduce the \emph{online} computation.

\textbf{Multi-server (information-theoretic) PIR.} The database is replicated across $t \geq 2$ non-colluding servers. Privacy holds information-theoretically: no single server (or coalition below the threshold) gains any information about $i$. The simplest 2-server scheme uses additive secret sharing of a selection vector, achieving $O(N)$ communication per server---already non-trivial compared to $O(NB)$ when $B$ is large.

\subsection{Distributed Point Functions (DPF)}
\label{sec:dpf}

A Distributed Point Function (DPF) \cite{gilboa2014dpf,boyle2015fss,boyle2016fss} is a cryptographic primitive that enables efficient 2-server PIR. A \emph{point function} $f_{\alpha,\beta}$ evaluates to $\beta$ at input $\alpha$ and to $0$ everywhere else. A DPF splits this point function into two keys $(k_0, k_1)$ such that:
\begin{itemize}[nolistsep,leftmargin=*]
    \item \textbf{Correctness:} $\text{Eval}(k_0, x) \oplus \text{Eval}(k_1, x) = f_{\alpha,\beta}(x)$ for all $x$.
    \item \textbf{Security:} Each individual key $k_b$ is computationally indistinguishable from a key for any other point function.
\end{itemize}

Each key has size $O(\lambda \log N)$ where $\lambda$ is the security parameter and $N$ is the domain size. To perform PIR on a database of $N$ entries indexed by $[N]$:
\begin{enumerate}[nolistsep,leftmargin=*]
    \item The client generates DPF keys $(k_0, k_1)$ for the point function $f_{i, 1}$ where $i$ is the target index.
    \item Server $b$ expands $k_b$ to obtain a vector of $N$ field elements and computes $r_b = \bigoplus_{x=0}^{N-1} \text{Eval}(k_b, x) \cdot D[x]$.
    \item The client recovers $D[i] = r_0 \oplus r_1$.
\end{enumerate}

The server computation is $O(NB)$ but consists of simple XOR operations over the database, making it extremely fast in practice---often close to the speed of a single linear pass over the data in memory.

\subsection{Cuckoo Hashing}

Cuckoo hashing \cite{PR01} is a hashing scheme that provides worst-case $O(1)$ lookup time. To store $n$ items, we maintain a table of $T$ slots organized into $\lceil T/s \rceil$ buckets of size $s$, with $h$ independent hash functions $H_1, \ldots, H_h$.

Each item $x$ can reside in any of its $h$ candidate buckets $H_1(x), \ldots, H_h(x)$. To insert $x$: if any candidate bucket has a free slot, place $x$ there. Otherwise, evict a random existing item from one candidate bucket, insert $x$, and re-insert the evicted item recursively. With $h \geq 3$ hash functions and bucket size $s \geq 4$, load factors above 95\% are achievable.

\textbf{PIR by keyword.} Cuckoo hashing enables PIR by keyword (as opposed to PIR by index). Given a keyword $w$, the client computes its $h$ candidate bucket indices and issues PIR queries for those buckets. Upon retrieving the bucket contents, the client checks which slot (if any) contains an entry matching $w$. This transforms the problem of searching an unordered key-value store into a constant number of PIR-by-index queries.

\subsection{Batch PIR and Probabilistic Batch Codes}

\textbf{Batch PIR} \cite{IKOS04} addresses the problem of retrieving $m$ entries simultaneously. The naive approach---issuing $m$ independent PIR queries---costs $m$ times the single-query cost. Batch PIR aims to reduce this by amortizing work across queries.

\textbf{Probabilistic Batch Codes (PBC)} achieve batch PIR by distributing $m$ queries across $K$ ``buckets'' (with $K > m$) such that each bucket contains at most one real query. The construction uses cuckoo hashing:
\begin{enumerate}[nolistsep,leftmargin=*]
    \item For each of the $m$ queries, compute $h$ candidate buckets using cuckoo hash functions.
    \item Place queries into buckets via cuckoo insertion with eviction chains.
    \item Pad all empty buckets with random dummy queries.
    \item Issue $K$ independent PIR queries in parallel---one per bucket.
\end{enumerate}

Since each bucket holds at most one real query, the server processes $K$ single-entry PIR queries. For DPF-PIR, this means generating $K$ DPF key pairs and having each server perform $K$ XOR-sum computations. The effective cost per query drops from $O(NB)$ to $O(NB \cdot K/m)$ when the server can share work across the $K$ evaluations.

\subsection{Bitcoin UTXO Model}

Bitcoin uses an Unspent Transaction Output (UTXO) model. Each transaction consumes previous outputs (inputs) and creates new outputs. The set of all outputs that have not yet been spent forms the \emph{UTXO set}, which represents the current state of all spendable coins.

Each UTXO is uniquely identified by a pair (TXID, vout), where TXID is the 32-byte hash of the creating transaction and vout is the output index within that transaction. Each UTXO carries an \emph{amount} (in satoshis) and a \emph{ScriptPubKey}---a locking script that specifies the conditions under which the output can be spent.

Common ScriptPubKey types include:
\begin{itemize}[nolistsep,leftmargin=*]
    \item \textbf{P2PKH} (Pay-to-Public-Key-Hash): Legacy addresses starting with \texttt{1}
    \item \textbf{P2SH} (Pay-to-Script-Hash): Script hash addresses starting with \texttt{3}
    \item \textbf{P2WPKH / P2WSH} (Pay-to-Witness-*): SegWit addresses starting with \texttt{bc1q}
    \item \textbf{P2TR} (Pay-to-Taproot): Taproot addresses starting with \texttt{bc1p}
\end{itemize}

A \emph{light client} knows its own private keys (and thus its addresses and ScriptPubKeys) but does not maintain a full copy of the blockchain. To spend its coins, the light client needs to discover which UTXOs are locked to its ScriptPubKeys and retrieve the corresponding (TXID, vout, amount) tuples. PIR enables this discovery without revealing the client's addresses to the server.

\clearpage

\section{PIR Protocol Landscape}
\label{sec:landscape}

We survey practical PIR constructions across two axes: the number of servers and the amount of client-side state. We consider four settings:
\begin{itemize}[nolistsep,leftmargin=*]
    \item \textbf{Single-server:}
    \smallskip
    \begin{itemize}[nolistsep,leftmargin=*]
        \item Almost no client storage: YPIR-over-SimplePIR \cite{ypir2024}

        \item Sublinear client storage without a linear pass: SimplePIR or DoublePIR \cite{henz23}

        \item Sublinear client storage after a linear pass: Plinko \cite{plinko2024}, RMS24 \cite{rms24}, or Piano \cite{piano2024}
    \end{itemize}

    \smallskip
    \item \textbf{2-server:}
    \smallskip
    \begin{itemize}[nolistsep,leftmargin=*]
        \item Almost no client storage: DPF-PIR \cite{gilboa2014dpf,boyle2015fss,boyle2016fss}

        \item Sublinear client storage: Plinko, RMS24, or Piano but moving the hint generation to a hint server
    \end{itemize}
\end{itemize}

\subsection{Single-Server PIR}

We focus on single-server PIR without per-client storage on the server and allow server to preprocess the database. For protocols that do not require client storage (or if the client storage is just for some data-independent parameters), YPIR-over-SimplePIR has achieved best performance and throughput for larger blocks, in comparison to Hintless PIR \cite{li24}, Tiptoe \cite{tiptoe23}, and Spiral \cite{spiral22}. 

If we allow the client to make data-dependent but sublinear storage for ``hints'', then we get better results with SimplePIR or DoublePIR, where the client downloads the hints before making any query. Note that DoublePIR is most suitable for single-bit or few-bit entries, and does not perform well when the entry size exceeds several tens of bits. This is because: (1) while the client hint is independent of the database size, it scales linearly with the entry size; and (2) this linear scaling has a very large multiplicative constant factor. This makes DoublePIR suitable for certain PIR use cases, but not the other. When using SimplePIR or DoublePIR, the server time appears to improve around $2$ times to an order of magnitudes with the sublinear hints over those that don't require hints.

If we do one step further and allow the client to download and make a linear pass over the database to generate sublinear hints (the client downloads and performs the linear pass in the streaming manner with only sublinear storage needed), then we can get much better performance with Plinko, RMS24, or Piano. The server time can be reduced by potentially two orders of magnitude. This setting works well for many devices with access to home internet or WiFi network where the network is not metered, but not all. 

\subsection{Two-Server PIR}

We now consider PIR that assumes two servers that are non-colluding and non-communicating, meaning that the two servers do not need to know each other, and it is only the user who engages with both servers. In fact, any two trustworthy servers that have the database can be selected by the user to run this protocol. This setting is especially relevant to Bitcoin peer-to-peer (P2P) network as a user can discover two full nodes that provide this capability and use their help. The caveat is that the user needs to trust that these two servers do not collude with each other. If not, the two colluding servers can immediately learn the user's query.

The easiest way to achieve 2-server PIR is to use DPF-PIR, which is based on distributed point functions (DPF) \cite{gilboa2014dpf,boyle2015fss,boyle2016fss}. Given a share of DPF keys of size $O(logN)$, each server can expand it to $N$ bits that look random, where the two servers' bits differ exactly on the location that the user queries. Each server retrieves the entries with bit $1$ and computes an XOR sum of them. The user finally performs an XOR over the two XOR sums from the two servers and obtains the desired query result. It requires $O(NB)$ server compute time with $B$ being the entry size, but the computation is extremely lightweight. Usually, the compute can answer a PIR query close to the time that it takes to do a linear pass of the database in the storage medium (which usually is the main memory).

If we allow hint generation, as is the case in the single-server setting, we know that the server can answer the PIR queries in sublinear time, which is extremely efficient that seems to ``surpass the memory speed.'' Previously in the single-server setting, a significant downside is that the client needs to download the entire database (in a streaming manner) to generate the hint locally. Now, in the 2-server setting, the user can outsource the hint generation to one of the servers (``hint server''), and the hint can be used to query any server other than the hint server. This avoids the need for the user to download the database, and the user can directly download the hint, though it still requires sublinear client storage.

\subsection{Query Volume Considerations}

When we compare these different solutions, it is important to understand how many queries a user would make. For example, if a user only makes one query, then all the compute and communication costs for the hint generation may not be worthwhile if it is single-use. The overhead is not just about the client, but also the server. For single-server protocols that require a linear pass, a server needs to serve the entire database to this user. For 2-server protocols with outsourced hint generation, the hint server bears the computation cost for doing that linear pass over its local copy of the database.

This is especially relevant to Bitcoin light client synchronization because a light client may maintain a small wallet that only cares about UTXOs that the wallet can spend. In this situation, the light client knows exactly how many queries it needs to make, and after those queries, the light client has no need to access this database anymore as it has been synced up to the state that is current to the database. In this case, the client may only make one or a few queries to the database, and therefore the overhead for hint generation may not be worthwhile. An example of such a use case is when a light client imports a seed phrase or a private key, in which case it needs to identify all the UTXOs associated with the imported wallets, but this is a one-time thing. 

Despite this concern, PIR is still superior than asking the light client to download the compact block filters for all the blocks since the imported key is first used (if known) and fetch a  number of blocks. First, the compact block filters for the entire Bitcoin history can still be a few gigabytes. Second, the usual way to fetch relevant blocks is to request these blocks directly from a full node by giving the block height, which leaks a little bit of privacy (which blocks this new wallet has a UTXO for). Whether this privacy leakage is acceptable depends on the situation and the user. Our goal building the PIR solution is to help those users who want to avoid this privacy leakage (which, as we will show, it is easy to avoid).

\subsection{Integrity and Verifiability}

Modern PIR naturally achieves privacy even against malicious servers, so we only need to focus on integrity. There are many ways to achieve integrity, some of which offer integrity against selective failure attacks \cite{kilian88,naor99,dwork99,camenisch07,lindell11,kiraz06}, while some do not. Whether it is okay to ignore selective failure attacks depends on the user and the use case. For example, a user who receives a potentially selectively failed response can continue to finish the rest of the protocol using random queries, pretending to have received a valid response, and then fetch the data from other servers.

If we ignore selective failure attacks, the most practical solution for single-server and 2-server solutions appears to be KZG polynomial commitment \cite{kzg10}, where the opening proofs are generated using the Feist-Khovratovich batch technique \cite{fk23}. An important limitation is that it uses elliptic curves, and therefore is not post-quantum. The KZG opening proof would add about 16 bytes to each entry. To reduce the PIR compute overhead caused by the additional opening proof, we can merge a few entries into one larger entry, and still only one opening proof is needed for the larger entry. With the same setup parameters of KZG, all the nodes in the network should have the same KZG polynomial commitment. 

This solution, however, fails to address selective failure attacks. Selective failure attacks refer to the case when a malicious server manipulates a subset $S$ of the database entries such that if the user is reading an entry in $S$, it would fail, and if the user is reading an entry outside of $S$, it would succeed. The concern is that such a selective failure, which may be realized by the malicious server through a change of the user's subsequent actions, can leak if the user is reading an entry in $S$.

To address selective failure attacks, we can choose protocols with this additional verifiability guarantee against selective failure attacks. For single-server protocols without a linear pass, RLP26 \cite{rlp26} is a good candidate, which improves over VeriSimplePIR \cite{cl24} and significantly reduces the client hint using some one-time computation. 
\begin{itemize}[nolistsep,leftmargin=*]
    \item \textbf{Single-server:}
    \smallskip
    \begin{itemize}[nolistsep,leftmargin=*]
        \item Almost no client storage: RLP26 \cite{rlp26}

        \item Sublinear client storage after a linear pass: Plinko, RMS24, or Piano with Merkle proofs, or verifiable BALANCED-PIR \cite{ar26}
    \end{itemize}

    \smallskip
    \item \textbf{2-server:} Plinko, RMS24, or Piano with Merkle proofs, or verifiable BALANCED-PIR with outsourced hint generation
\end{itemize}

\clearpage

\section{PIR for UTXO Set}
\label{sec:utxo-pir}

A light client knowing a receiving ScriptPubKey can use PIR to pull the full 32-byte TXID, vout, and amount for each UTXO under that ScriptPubKey---all the information needed to spend. Our system uses a two-level architecture: an \emph{index map} that locates data, and a \emph{chunk map} that stores the actual UTXO data. Full 32-byte TXIDs are stored directly in the chunks, eliminating a separate TXID resolution step.

\begin{figure}[htbp]
\centering
\begin{tikzpicture}[
    node distance=1.2cm and 1.8cm,
    box/.style={rectangle, draw, rounded corners, minimum width=2.5cm, minimum height=1cm, align=center, font=\small},
    data/.style={rectangle, draw, fill=blue!10, minimum width=1.8cm, minimum height=0.7cm, align=center, font=\footnotesize},
    arrow/.style={-{Stealth[length=2.5mm]}, thick},
    label/.style={font=\scriptsize, text=gray!70!black},
    pirlabel/.style={font=\scriptsize\bfseries, text=blue!60!black}
]

\node[data, fill=green!15] (user) {User's\\ScriptPubKey(s)};
\node[box, fill=orange!20, right=2.5cm of user] (map1) {Index Map\\{\tiny ($K{=}75$ batch PIR)}};
\node[data, fill=yellow!20, right=of map1] (chunkinfo) {Chunk ID\\+ Data Length};
\node[box, fill=orange!20, below=1cm of map1] (map2) {Chunk Map\\{\tiny ($K_\text{chunk}{=}80$ batch PIR)}};
\node[data, fill=yellow!20, right=of map2] (utxo) {UTXO Data\\{\tiny (32B TXID, vout, amount)}};

\draw[arrow] (user) -- (map1) node[midway, above, pirlabel] {PBC + Cuckoo};
\draw[arrow] (map1) -- (chunkinfo) node[midway, above, label] {$K$ DPF queries};
\draw[arrow] (chunkinfo.south) -- ++(0,-0.3) -| (map2.north);
\draw[arrow] (map2) -- (utxo) node[midway, above, label] {$K_\text{chunk}$ DPF queries};

\begin{scope}[on background layer]
    \node[fit=(map1)(chunkinfo), draw=gray!50, dashed, rounded corners, inner sep=8pt, label={[font=\tiny\itshape, text=gray]above:Level 1}] {};
    \node[fit=(map2)(utxo), draw=gray!50, dashed, rounded corners, inner sep=8pt, label={[font=\tiny\itshape, text=gray]left:Level 2}] {};
\end{scope}

\end{tikzpicture}
\caption{Two-level batch PIR architecture. The client issues $K$ parallel DPF-PIR queries per level, using Probabilistic Batch Codes to pack multiple lookups into a single round.}
\label{fig:two-maps}
\end{figure}

\textbf{Sort and group.} The server sorts UTXOs by ScriptPubKey hash, then groups them into fixed-size chunks. Each chunk stores complete UTXO records including full 32-byte TXIDs, vout indices, and amounts. Multiple ScriptPubKeys may share a chunk via bin-packing.

\textbf{Index map.} Maps each ScriptPubKey (via its RIPEMD-160 hash) to the chunk location and data length. This is organized as a cuckoo hash table for $O(1)$ keyword-to-index lookup \cite{PR01}. Each index entry contains: the 20-byte script hash (for verification that the query hit a real entry), a chunk identifier, and the byte length of the UTXO data within that chunk.

\textbf{Batch PIR via Probabilistic Batch Codes (PBC).} To query multiple entries simultaneously, we use PBC \cite{IKOS04} implemented via cuckoo hashing. For $m$ queries, the client computes 3 candidate buckets per query using cuckoo hash functions, then places queries into $K$ buckets using cuckoo insertion (with eviction chains up to 500 kicks). Unfilled buckets receive random dummy queries. All $K$ DPF-PIR queries are issued in parallel to both servers, achieving an effective throughput improvement proportional to $K/m$.

\textbf{Distributional PIR.} We can apply distributional PIR \cite{lehmkuhl25} to optimize for the common case: most light clients hold few UTXOs under common script types, while whale addresses with thousands of UTXOs likely run full nodes. By tuning the PIR parameters for the more likely queries, the average server cost decreases without eliminating the ability to query less common entries.

\clearpage

\section{Implementation}
\label{sec:implementation}

We have implemented a complete system based on the 2-server DPF-PIR approach with Probabilistic Batch Codes (PBC), as described in Section~\ref{sec:utxo-pir}. The implementation consists of database generation tools (Rust), two PIR servers (Rust with \texttt{tokio}), and client libraries (Rust CLI and TypeScript/Web). The two-level architecture from Figure~\ref{fig:two-maps} eliminates the TXID resolution step of earlier designs by storing full 32-byte TXIDs directly in the UTXO chunks.

\subsection{System Architecture}

Each server maintains identical copies of the index and chunk databases, loaded via memory-mapped files (\texttt{memmap2}). The client connects to both servers over WebSocket and distributes DPF key shares, so neither server learns the query target unless they collude. Communication uses a compact binary protocol supporting database discovery, batch PIR queries, and connection health checks.

\begin{figure}[htbp]
\centering
\begin{tikzpicture}[
    node distance=1.5cm and 2cm,
    box/.style={rectangle, draw, rounded corners, minimum width=3cm, minimum height=1.2cm, align=center, font=\small},
    db/.style={cylinder, draw, shape border rotate=90, aspect=0.25, minimum height=1.5cm, minimum width=1.2cm, fill=blue!10, font=\footnotesize},
    arrow/.style={-{Stealth[length=2.5mm]}, thick},
    label/.style={font=\scriptsize, text=gray!70!black}
]

\node[box, fill=green!20] (client) {Client\\{\scriptsize (Rust CLI / Web)}};
\node[box, fill=orange!20, below left=1.5cm and 1cm of client] (server1) {Server 1};
\node[box, fill=orange!20, below right=1.5cm and 1cm of client] (server2) {Server 2};
\node[db, below=0.5cm of server1] (db1) {Index + Chunks};
\node[db, below=0.5cm of server2] (db2) {Index + Chunks};

\draw[arrow] (client) -- (server1) node[midway, above, sloped, label] {$K$ DPF keys $\{k_1^{(i)}\}$};
\draw[arrow] (client) -- (server2) node[midway, above, sloped, label] {$K$ DPF keys $\{k_2^{(i)}\}$};
\draw[arrow, dashed, gray] (server1) -- (db1);
\draw[arrow, dashed, gray] (server2) -- (db2);

\draw[arrow, blue!60!black] (server1.north east) -- ++(0.3, 0.5) node[right, font=\scriptsize] {$\{r_1^{(i)}\}$};
\draw[arrow, blue!60!black] (server2.north west) -- ++(-0.3, 0.5) node[left, font=\scriptsize] {$\{r_2^{(i)}\}$};

\node[above=1.8cm of client, font=\small, text=blue!60!black] (xor) {Results: $r_1^{(i)} \oplus r_2^{(i)}$ for $i \in [K]$};
\draw[arrow, blue!60!black, dashed] ($(server1.north east)+(0.3, 0.5)$) -- (xor);
\draw[arrow, blue!60!black, dashed] ($(server2.north west)+(-0.3, 0.5)$) -- (xor);

\end{tikzpicture}
\caption{Two-server batch DPF-PIR architecture. The client generates $K$ DPF key pairs (one per PBC bucket) and sends the shares to each server. Each server computes $K$ XOR sums in parallel; the client combines responses to recover all queried entries.}
\label{fig:architecture}
\end{figure}

\subsection{Database Generation}

The offline pipeline transforms raw Bitcoin blockchain data into two PIR-queryable databases. Each stage is a separate Rust binary in \texttt{build\_db/src/bin}.

\textbf{Pipeline overview:}
\begin{enumerate}[nolistsep,leftmargin=*]
    \item \textbf{UTXO extraction:} Parse the UTXO snapshot (\texttt{dumptxoutset}) to extract ScriptPubKey hashes, full 32-byte TXIDs, vout indices, and amounts.
    \item \textbf{Chunk generation:} Group UTXOs by ScriptPubKey, serialize with VarInt encoding, and bin-pack into fixed-size chunks (default 256 bytes). Each chunk stores complete records including full TXIDs.
    \item \textbf{Index construction:} Build an index mapping each ScriptPubKey hash (20 bytes) to its chunk location and data length.
    \item \textbf{Cuckoo hashing:} Organize both the index and chunk databases into cuckoo hash tables (bucket size 4, 3 hash functions) for PBC-compatible batch retrieval.
    \item \textbf{Verification:} Validate that all entries are retrievable from the cuckoo tables.
\end{enumerate}

\textbf{Dual role of cuckoo hashing.} Cuckoo hashing serves two purposes in our system. First, it provides PIR-by-keyword: given a ScriptPubKey, the client computes candidate bucket indices to locate data without scanning. Second, it implements PBC for batch PIR: when querying $m$ entries, each query maps to 3 candidate buckets; the client places queries into $K$ total buckets via cuckoo insertion, and all $K$ DPF queries execute in parallel.

\begin{table}[htbp]
\centering
\begin{tabular}{|l|c|c|c|}
\hline
\textbf{Database} & \textbf{Entry Size} & \textbf{Buckets ($K$)} & \textbf{DPF domain ($N$)} \\
\hline
Index (Level 1) & 28 B & 75 & $2^{20}$ \\
Chunks (Level 2) & 256 B & 80 & $2^{20}$ \\
\hline
\end{tabular}
\caption{Batch PIR parameters. Each level uses $K$ parallel DPF-PIR queries over tables of $N$ entries.}
\label{tab:databases}
\end{table}

\subsection{Query Protocol}

A complete UTXO lookup requires two rounds of batch PIR:

\textbf{Level 1: Index lookup.} The client computes RIPEMD-160 hashes of all queried ScriptPubKeys, derives 3 candidate buckets per query via cuckoo hash functions (\texttt{splitmix64}-based), and places them into $K{=}75$ PBC buckets. For each bucket, the client generates a DPF key pair targeting the assigned row in the index table and sends one share to each server. The servers expand all $K$ DPF keys and return $K$ XOR sums. The client combines responses to recover index entries containing chunk IDs and data lengths.

\textbf{Level 2: Chunk retrieval.} Using the chunk locations from Level~1, the client repeats the process with $K_\text{chunk}{=}80$ PBC buckets over the chunk database. The recovered chunks contain full UTXO records: 32-byte TXIDs, vout indices, and amounts in satoshis---everything needed to construct a spending transaction.

\subsection{Client Implementations}

\textbf{Rust CLI (\texttt{lookup\_pir}).} Command-line tool that takes ScriptPubKeys as hex, performs the two-level batch PIR protocol, and outputs decoded UTXO data.

\textbf{Web client (TypeScript).} Browser-compatible library using WebSocket connections. Implements DPF key generation, cuckoo hashing, and PBC placement in pure TypeScript. Supports Bitcoin address input (P2PKH, P2SH, P2WPKH, P2WSH, P2TR) with automatic conversion to ScriptPubKey. The web interface provides a self-contained demo at \texttt{batch\_web\_client/}.

\clearpage

\section{Extension: Further Reducing the UTXO Set}
\label{sec:extension}

Since the PIR service is designed for light clients who do not frequently run the Bitcoin node for synchronization or for newly imported wallet addresses from a regular user, we can realistically expect that the user only makes a few transactions, keeps only a few UTXOs in the chain, and has no intention to create large number of dust outputs (a typical pattern of some Bitcoin layer-2 protocols). Therefore, taking this in mind, we can reduce the UTXO set by filtering out scriptPubKeys that show different user patterns, or, if not filtering out the entire scriptPubKey, we can selectively filter out certain transactions or certain UTXOs that match some profiles.

For example, a scriptPubKey that has been involved in a significantly above average number of UTXOs is likely not a regular user and probably should run its own full node to have the full security and privacy. 

Another example is that we can filter out UTXOs that are of dust amount. It is possible that a honest user has dust UTXOs and may want to use them in the future, but given that each dust UTXO is worth probably like $0.4$ USD, we can say that they are of a lower prioirity, and a user who sincerely wants to use them should first try to avoid generating dust UTXOs as it should not occur naturally, and if the user sincerely wants to recover from the dust amount, the user is recommended to use an RPC server (maybe through Tor), or if strict privacy is a concern, run a full node---honestly speaking, if the user wants to consolidate these dust UTXOs, the other full nodes are going to learn it sooner or later anyway because they can wait for the consolidation transaction. 

If we worry that the dust amount filtering might be excessive, we can also target certain protocols that are known to generate a large number of dust UTXOs, such as the Omnilayer, by looking into the \textrm{OP\_RETURN} output next to the dust UTXO and checking if it starts with ``omni''.

We can also make a list of known exchange wallet addresses and filter them out given that these crypto exchange is definitely not running a light client. There are some services for this purpose, such as Arkham, that provides a list of identified addresses of notable market paricipants (who are unlikely using this light client).

\clearpage

\section*{Acknowledgments}

We thank \href{https://wuwuz.github.io/}{Mingxun Zhou} for advice and suggestions on stateful PIR and their implementations. We thank \href{https://sites.google.com/view/renling}{Ling Ren} and \href{https://scholar.google.com/citations?user=f54M0ikAAAAJ&hl=en}{Pranav Shriram Arunachalaramanan} for advice and suggestions on state-of-the-art PIR schemes in various settings, batch PIR, and understanding of the SimplePIR family of protocols.

\bibliography{refs}

\end{document}